---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->
[![codecov.io](https://codecov.io/github/darwin-eu/CodelistGenerator/coverage.svg?branch=main)](https://codecov.io/github/darwin-eu/CodelistGenerator?branch=main)
[![R-CMD-check](https://github.com/darwin-eu/CodelistGenerator/workflows/R-CMD-check/badge.svg)](https://github.com/darwin-eu/CodelistGenerator/actions)
[![Lifecycle:Experimental](https://img.shields.io/badge/Lifecycle-Experimental-339999)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

# CodelistGenerator

## Introduction
CodelistGenerator is used to create a candidate set of codes for helping to define patient cohorts in data mapped to the OMOP common data model. A little like the process for a systematic review, the idea is that for a specified search strategy, CodelistGenerator will identify a set of concepts that may be relevant, with these then being screened to remove any irrelevant codes.

## Installation
You can install the development version of CodelistGenerator like so:
``` r
install.packages("remotes")
remotes::install_github("darwin-eu/CodelistGenerator")
```

## Connecting to the OMOP CDM vocabularies
### Option 1: Connect to a live OMOP CDM database
```{r, eval=FALSE}
# example with postgres database connection details
server_dbi<-Sys.getenv("server")
user<-Sys.getenv("user")
password<- Sys.getenv("password")
port<-Sys.getenv("port")
host<-Sys.getenv("host")

db <- DBI::dbConnect(RPostgres::Postgres(),
                dbname = server_dbi,
                port = port,
                host = host,
                user = user,
                password = password)

# name of vocabulary schema
vocabulary_database_schema<-Sys.getenv("vocabulary_schema")
```

### Option 2: Download the vocabularies from Athena 
You will first need to obtain the OMOP CDM vocabularies from https://athena.ohdsi.org. Once these are downloaded, you can make a vocabulary only SQLite database like so:
```{r, eval=FALSE}
vocab.folder<-Sys.getenv("omop_cdm_vocab_path") # path to directory of unzipped files
concept<-read_delim(paste0(vocab.folder,"/CONCEPT.csv"),
     "\t", escape_double = FALSE, trim_ws = TRUE)
concept_relationship<-read_delim(paste0(vocab.folder,"/CONCEPT_RELATIONSHIP.csv"),
     "\t", escape_double = FALSE, trim_ws = TRUE) 
concept_ancestor<-read_delim(paste0(vocab.folder,"/CONCEPT_ANCESTOR.csv"),
     "\t", escape_double = FALSE, trim_ws = TRUE)
concept_synonym<-read_delim(paste0(vocab.folder,"/CONCEPT_SYNONYM.csv"),
     "\t", escape_double = FALSE, trim_ws = TRUE)
vocabulary<-read_delim(paste0(vocab.folder,"/VOCABULARY.csv"),
     "\t", escape_double = FALSE, trim_ws = TRUE)

db <- dbConnect(RSQLite::SQLite(), ":memory:")
dbWriteTable(db, "concept", concept, overwrite=TRUE)
dbWriteTable(db, "concept_relationship", concept_relationship, overwrite=TRUE)
dbWriteTable(db, "concept_ancestor", concept_ancestor, overwrite=TRUE)
dbWriteTable(db, "concept_synonym", concept_synonym, overwrite=TRUE)
dbWriteTable(db, "vocabulary", vocabulary)
rm(concept,concept_relationship, concept_ancestor, concept_synonym, vocabulary)

vocabulary_database_schema<-"main"
```

## Example search
Every codelist is specific to a version of the OMOP CDM vocabularies, so we can first check the version.
```{r example,  message=FALSE, warning=FALSE,echo=FALSE}
library(DBI)
library(RPostgres)
library(dplyr)
library(CodelistGenerator)

# usethis::edit_r_environ()
server_dbi<-Sys.getenv("SERVER_DBI_FEB22")
user<-Sys.getenv("DB_USER_FEB22")
password<- Sys.getenv("DB_PASSWORD_FEB22")
port<-Sys.getenv("DB_PORT_FEB22")
host<-Sys.getenv("DB_HOST_FEB22")

db <- dbConnect(RPostgres::Postgres(),
                dbname = server_dbi,
                port = port,
                host = host,
                user = user,
                password = password)
vocabulary_database_schema<-Sys.getenv("DB_VOCAB_FEB22")
```


```{r,  message=FALSE, warning=FALSE}
dplyr::tbl(db, dplyr::sql(paste0(
    "SELECT * FROM ",
    vocabulary_database_schema,
    ".vocabulary"
    ))) %>%
    dplyr::rename_with(tolower) %>%
    dplyr::filter(.data$vocabulary_id == "None") %>%
    dplyr::select("vocabulary_version") %>%
    dplyr::collect() %>%
    dplyr::pull()
```

We can then search for asthma like so
```{r,  message=FALSE, warning=FALSE}
asthma_1<-get_candidate_codes(keywords="asthma",
                    domains = "Condition",
                    db=db,
                    vocabulary_database_schema = vocabulary_database_schema)
head(asthma_1, 10)
```

Perhaps we want to exclude certain concepts as part of the search strategy, in which case this can be added like so
```{r,  message=FALSE, warning=FALSE}
asthma_2<-get_candidate_codes(keywords="asthma",
                    domains = "Condition",
                    exclude = "Poisoning by antiasthmatic",
                    db=db,
                    vocabulary_database_schema = vocabulary_database_schema)
head(asthma_2, 10)
```

We can then also see source codes these are mapped from, for example
```{r,  message=FALSE, warning=FALSE}
asthma_icd_mappings<-show_mappings(candidate_codelist=asthma_2,
                     source_vocabularies="ICD10CM",
                    db=db,
                    vocabulary_database_schema =  vocabulary_database_schema)
head(asthma_icd_mappings %>% 
       select("Standard concept_id name",
              "Source name" ),
     10)

dbDisconnect(db)
```
