---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# CodelistGenerator

## Introduction
CodelistGenerator is used to create a candidate set of codes for helping to define patient cohorts in data mapped to the OMOP common data model. A little like the process for a systematic review, the idea is that for a specified search strategy, CodelistGenerator will identify a set of concepts that may be relevant, with these then being screened to remove any irrelevant codes.

## Installation
You can install the development version of CodelistGenerator like so:
``` r
install.packages("remotes")
remotes::install_github("oxford-pharmacoepi/CodelistGenerator")
```

## Example
Note, Eunomia, used in the example below, does not include a full set of vocabularies. 
```{r,  message=FALSE, warning=FALSE}
library(CodelistGenerator)
library(dplyr)
library(Eunomia)
library(RSQLite)
library(DBI)
untar(xzfile(system.file("sqlite", "cdm.tar.xz", package = "Eunomia"), open = "rb"),
        exdir =  tempdir())
db <- dbConnect(RSQLite::SQLite(), paste0(tempdir(),"\\cdm.sqlite"))

```

Every codelist is specific to a version of the OMOP CDM vocabularies, so we can first check the version.
```{r example,  message=FALSE, warning=FALSE}
get_vocab_version(db=db,
                  vocabulary_database_schema = "main")
```

We can then search for asthma like so
```{r,  message=FALSE, warning=FALSE}
get_candidate_codes(keywords="asthma",
                    domains = "Condition",
                    db=db,
                    vocabulary_database_schema = "main")
```

Perhaps we want to exclude asthma in children as part of the search strategy, in which case this can be added like so
```{r,  message=FALSE, warning=FALSE}
get_candidate_codes(keywords="asthma",
                    domains = "Condition",
                    exclude = "Childhood asthma",
                    db=db,
                    vocabulary_database_schema = "main")

```
