---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->
[![codecov.io](https://codecov.io/github/darwin-eu/CodelistGenerator/coverage.svg?branch=main)](https://codecov.io/github/darwin-eu/CodelistGenerator?branch=main)
[![R-CMD-check](https://github.com/darwin-eu/CodelistGenerator/workflows/R-CMD-check/badge.svg)](https://github.com/darwin-eu/CodelistGenerator/actions)
[![Lifecycle:Experimental](https://img.shields.io/badge/Lifecycle-Experimental-339999)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

# CodelistGenerator

## Introduction
CodelistGenerator is used to create a candidate set of codes for helping to define patient cohorts in data mapped to the OMOP common data model. A little like the process for a systematic review, the idea is that for a specified search strategy, CodelistGenerator will identify a set of concepts that may be relevant, with these then being screened to remove any irrelevant codes.

## Installation
You can install the development version of CodelistGenerator like so:
``` r
install.packages("remotes")
remotes::install_github("darwin-eu/CodelistGenerator")
```

## Connecting to the OMOP CDM vocabularies
```{r, eval=FALSE}
# First load required libraries
library(DBI)
library(dplyr)
library(dbplyr)
library(CodelistGenerator)
library(CDMConnector)
# Note that you will also need another library, like RPostgres, to make your database connection
```

```{r, eval=FALSE}
# example with postgres database connection details
serverDbi <- Sys.getenv("server")
user <- Sys.getenv("user")
password <- Sys.getenv("password")
port <- Sys.getenv("port")
host <- Sys.getenv("host")

db <- DBI::dbConnect(RPostgres::Postgres(),
  dbname = serverDbi,
  port = port,
  host = host,
  user = user,
  password = password
)

# name of vocabulary schema
vocabularyDatabaseSchema <- Sys.getenv("vocabulary_schema")
```

```{r,  message=FALSE, warning=FALSE, echo=FALSE}
library(DBI)
library(RPostgres)
library(dplyr)
library(dbplyr)
library(CodelistGenerator)
library(kableExtra)
library(CDMConnector)

# usethis::edit_r_environ()
serverDbi <- Sys.getenv("serverDbi_FEB22")
user <- Sys.getenv("DB_USER_FEB22")
password <- Sys.getenv("DB_PASSWORD_FEB22")
port <- Sys.getenv("DB_PORT_FEB22")
host <- Sys.getenv("DB_HOST_FEB22")

db <- dbConnect(RPostgres::Postgres(),
  dbname = serverDbi,
  port = port,
  host = host,
  user = user,
  password = password
)
vocabularyDatabaseSchema <- Sys.getenv("DB_VOCAB_FEB22")
```

## Create cdm_reference 
We first set up a `cdm_reference` using the CDMConncetor package. 

```{r,  message=FALSE, warning=FALSE}
cdm <- cdm_from_con(
  con = db,
  cdm_schema = vocabularyDatabaseSchema,
  cdm_tables = tidyselect::all_of(c(
    "concept",
    "concept_relationship",
    "concept_ancestor",
    "concept_synonym",
    "vocabulary"
  ))
)
```

## Example search
Every codelist is specific to a version of the OMOP CDM vocabularies, so we can first check the version.

```{r,  message=FALSE, warning=FALSE}
getVocabVersion(cdm = cdm)
```

We can then search for asthma like so
```{r,  message=FALSE, warning=FALSE, eval=FALSE}
asthma1 <- getCandidateCodes(
  cdm = cdm,
  keywords = "asthma",
  domains = "Condition"
)
head(asthma1, 10)
```

```{r,  message=FALSE, warning=FALSE, echo=FALSE}
asthma1 <- getCandidateCodes(
  cdm = cdm,
  keywords = "asthma",
  domains = "Condition"
)
kable(head(asthma1, 10))
```

Perhaps we want to exclude certain concepts as part of the search strategy, in which case this can be added like so
```{r,  message=FALSE, warning=FALSE, eval=FALSE}
asthma2 <- getCandidateCodes(
  keywords = "asthma",
  domains = "Condition",
  exclude = "Poisoning by antiasthmatic",
  cdm = cdm
)
head(asthma2, 10)
```

```{r,  message=FALSE, warning=FALSE, echo=FALSE}
asthma2 <- getCandidateCodes(
  cdm = cdm,
  keywords = "asthma",
  domains = "Condition",
  exclude = "Poisoning by antiasthmatic"
)
kable(head(asthma2, 10))
```

We can then also see non-standard codes these are mapped from, for example
```{r,  message=FALSE, warning=FALSE, eval=FALSE}
asthmaIcdMappings <- getMappings(
  cdm = cdm,
  candidateCodelist = asthma2,
  nonStandardVocabularies = "ICD10CM"
)
head(
  asthmaIcdMappings %>%
    select(
      standard_concept_name,
      standard_vocabulary_id,
      non_standard_concept_name,
      non_standard_vocabulary_id
    ),
  10
)
```

```{r,  message=FALSE, warning=FALSE, echo=FALSE}
asthmaIcdMappings <- getMappings(
  cdm = cdm,
  candidateCodelist = asthma2,
  nonStandardVocabularies = "ICD10CM"
)
kable(head(
  asthmaIcdMappings %>%
    select(
      standard_concept_name,
      standard_vocabulary_id,
      non_standard_concept_name,
      non_standard_vocabulary_id
    ),
  10
))
```


```{r,  message=FALSE, warning=FALSE, echo=FALSE}
dbDisconnect(db)
```
