---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  message=FALSE, 
  warning=FALSE 
)
```

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/CodelistGenerator)](https://CRAN.R-project.org/package=CodelistGenerator)
[![codecov.io](https://codecov.io/github/darwin-eu/CodelistGenerator/coverage.svg?branch=main)](https://codecov.io/github/darwin-eu/CodelistGenerator?branch=main)
[![R-CMD-check](https://github.com/darwin-eu/CodelistGenerator/workflows/R-CMD-check/badge.svg)](https://github.com/darwin-eu/CodelistGenerator/actions)
[![Lifecycle:Stable](https://img.shields.io/badge/Lifecycle-Stable-97ca00)](https://lifecycle.r-lib.org/articles/stages.html)
<!-- badges: end -->

# CodelistGenerator

## Introduction
CodelistGenerator is used to create a candidate set of codes for helping to define patient cohorts in data mapped to the OMOP common data model. A little like the process for a systematic review, the idea is that for a specified search strategy, CodelistGenerator will identify a set of concepts that may be relevant, with these then being screened to remove any irrelevant codes.

## Installation
You can install the development version of CodelistGenerator like so:
``` r
install.packages("remotes")
remotes::install_github("darwin-eu/CodelistGenerator")
```

## Example usage
```{r}
library(dplyr)
library(CDMConnector)
library(CodelistGenerator)
library(kableExtra)
```

In this example we'll use the Eunomia dataset (which only contains a subset of the OMOP CDM vocabularies)
```{r}
db <- DBI::dbConnect(duckdb::duckdb(), dbdir = eunomia_dir())
cdm <- cdm_from_con(db,
  cdm_schema = "main",
  cdm_tables = tidyselect::all_of(c(
    "concept",
    "concept_relationship",
    "concept_ancestor",
    "concept_synonym",
    "vocabulary"
  ))
)
```

Although we can run the search using vocabulary tables in the database or loaded into R, the fastest approach is using arrow. So let's create a new cdm reference using arrow (in this example saved to the temp directory, but in practice you could of course save files elsewhere for reuse).

```{r}
# save cdm vocabulary tables to temp directory
dOut<-here::here(tempdir(), "db_vocab")
dir.create(dOut)
CDMConnector::stow(cdm, dOut)
# new cdm reference using arrow
cdm_arrow <- CDMConnector::cdm_from_files(path = dOut, 
                                          as_data_frame = FALSE)
```

Every code list is specific to a version of the OMOP CDM vocabularies, so we can first check the version for Eunomia.
```{r}
getVocabVersion(cdm = cdm_arrow)
```

We can then search for asthma like so
```{r}
getCandidateCodes(
  cdm = cdm_arrow,
  keywords = "asthma",
  domains = "Condition"
) %>% 
  glimpse()
```

Perhaps we want to exclude certain concepts as part of the search strategy, in which case this can be added like so
```{r}
getCandidateCodes(
  cdm = cdm_arrow,
  keywords = "asthma",
  exclude = "childhood",
  domains = "Condition"
) %>% 
  glimpse()
```

We can then also see non-standard codes these are mapped from, for example here we can see the non-standard ICD10 code that maps to a standard snomed code for gastrointestinal hemorrhage returned by our search
```{r}
Gastrointestinal_hemorrhage <- getCandidateCodes(
  cdm = cdm_arrow,
  keywords = "Gastrointestinal hemorrhage",
  domains = "Condition"
)
Gastrointestinal_hemorrhage %>% 
  glimpse()
```

```{r}
getMappings(
  cdm = cdm_arrow,
  candidateCodelist = Gastrointestinal_hemorrhage,
  nonStandardVocabularies = "ICD10CM"
) %>% 
  glimpse()

```

```{r, echo=FALSE}
DBI::dbDisconnect(db)
```
