---
title: "Introduction to CodelistGenerator"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{01_Introduction_to_CodelistGenerator}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Creating a code list for dementia
For this example we are going to generate a candidate codelist for dementia, only looking for codes in the condition domain. Let's first load some libraries

```{r,  message=FALSE, warning=FALSE,echo=FALSE}
library(here)
library(readr)
library(DBI)
library(here)
library(dplyr)
library(stringr)
library(DT)
library(kableExtra)
library(CodelistGenerator)
library(CDMConnector)
```

## Connect to the OMOP CDM vocabularies
CodelistGenerator works with a `cdm_reference` to the vocabularies tables of the OMOP CDM using the CDMConnector package. This `cdm_reference` could be to the tables in a database, using an arrow_table, or loaded into R. CodelistGenerator will generally work fastest with an arrow_table so let's first set that up.

```{r, eval=FALSE}
# example with postgres database connection details
db <- DBI::dbConnect(RPostgres::Postgres(),
  dbname = Sys.getenv("server"),
  port = Sys.getenv("port"),
  host = Sys.getenv("host"),
  user = Sys.getenv("user"),
  password = Sys.getenv("password")
)
# name of vocabulary schema
vocabularyDatabaseSchema <- Sys.getenv("vocabulary_schema")

# create cdm reference
cdm <- CDMConnector::cdm_from_con(
  con = db,
  cdm_schema = vocabularyDatabaseSchema,
  cdm_tables = tidyselect::all_of(c(
    "concept",
    "concept_relationship",
    "concept_ancestor",
    "concept_synonym",
    "vocabulary"
  ))
)
# vocab to arrow
# save in temp folder for this example
dOut <- here(tempdir(), "db_vocab")
dir.create(dOut)
CDMConnector::stow(cdm, dOut)

# new cdm reference using arrow
cdmArrow <- CDMConnector::cdm_from_files(
  path = dOut,
  cdm_tables = tidyselect::all_of(c(
    "concept",
    "concept_relationship",
    "concept_ancestor",
    "concept_synonym",
    "vocabulary"
  )),
  as_data_frame = FALSE
)

rm(cdm)
```


## Check version of the vocabularies
It is important to note that the results from CodelistGenerator will be specific to a particular version of the OMOP CDM vocabularies. We can see the version of the vocabulary being used like so
```{r, eval=FALSE}
getVocabVersion(cdm = cdmArrow)
```

```{r,  message=FALSE, warning=FALSE,echo=FALSE}
vocabVersion <- readRDS(here("vignettes", "introVocab.RData"))
vocabVersion
```

## A code list from "Dementia" (4182210) and its descendants
The simplest approach to identifying potential codes is to take a high-level code and include all its descendants. 


```{r, eval=FALSE}
codesFromDescendants <- tbl(
  db,
  sql(paste0(
    "SELECT * FROM ",
    vocabularyDatabaseSchema,
    ".concept_ancestor"
  ))
) %>%
  filter(ancestor_concept_id == "4182210") %>%
  select("descendant_concept_id") %>%
  rename("concept_id" = "descendant_concept_id") %>%
  left_join(tbl(db, sql(paste0(
    "SELECT * FROM ",
    vocabularyDatabaseSchema,
    ".concept"
  )))) %>%
  select(
    "concept_id", "concept_name",
    "domain_id", "vocabulary_id"
  ) %>%
  collect()
```

```{r,  message=FALSE, warning=FALSE,echo=FALSE}
codesFromDescendants <- readRDS(here("vignettes", "introData01.RData"))
```

```{r,  message=FALSE, warning=FALSE }  
datatable(codesFromDescendants,
  rownames = FALSE,
  options = list(
    pageLength = 10,
    lengthMenu = c(10, 20, 50)
  )
)
```

This looks to pick up most relevant codes. But, this approach misses codes that are not a descendant of 4182210. For example, codes such as "Wandering due to dementia" (37312577; https://athena.ohdsi.org/search-terms/terms/37312577) and "Anxiety due to dementia" (37312031; https://athena.ohdsi.org/search-terms/terms/37312031) are not picked up.

## Generating a candidate code list using CodelistGenerator
To try and include all such terms that could be included we can use CodelistGenerator.

First, let's do a simple search for a single keyword of "dementia", including descendants of the identified codes.
```{r, eval=FALSE }
dementiaCodes1 <- getCandidateCodes(
  cdm = cdm,
  keywords = "dementia",
  domains = "Condition",
  includeDescendants = TRUE
)
```

```{r,  message=FALSE, warning=FALSE,echo=FALSE}
dementiaCodes1 <- readRDS(here("vignettes", "introData02.RData"))
```

```{r,  message=FALSE, warning=FALSE }
datatable(dementiaCodes1,
  rownames = FALSE,
  options = list(
    pageLength = 10,
    lengthMenu = c(10, 20, 50)
  )
)
```


## Comparing code lists
What is the difference between this code list and the one from 4182210 and its descendants?
```{r,  eval=FALSE }
codeComparison <- compareCodelists(
  codesFromDescendants,
  dementiaCodes1
)
```

```{r,  message=FALSE, warning=FALSE,echo=FALSE}
codeComparison <- readRDS(here("vignettes", "introData03.RData"))
```

```{r,  message=FALSE, warning=FALSE } 
kable(codeComparison %>%
  group_by(codelist) %>%
  tally())
```

What are these extra codes picked up by CodelistGenerator?
```{r,  message=FALSE, warning=FALSE }
datatable(
  codeComparison %>%
    filter(codelist == "Only codelist 2"),
  rownames = FALSE,
  options = list(
    pageLength = 10,
    lengthMenu = c(10, 20, 50)
  )
)
```

## Review mappings from non-standard vocabularies
Perhaps we want to see what ICD10CM codes map to our candidate code list. We can get these by running

```{r,  message=FALSE, warning=FALSE,echo=FALSE}
icdMappings <- readRDS(here("vignettes", "introData04.RData"))
```

```{r,  eval=FALSE }
icdMappings <- getMappings(
  cdm = cdm,
  candidateCodelist = dementiaCodes1,
  nonStandardVocabularies = "ICD10CM"
)
```

```{r,  message=FALSE, warning=FALSE }
datatable(icdMappings,
  rownames = FALSE,
  options = list(
    pageLength = 10,
    lengthMenu = c(10, 20, 50)
  )
)
```


```{r,  message=FALSE, warning=FALSE,echo=FALSE}
readMappings <- readRDS(here("vignettes", "introData05.RData"))
```

```{r,  eval=FALSE }
readMappings <- getMappings(
  cdm = cdm,
  candidateCodelist = dementiaCodes1,
  nonStandardVocabularies = "Read"
)
```

```{r,  message=FALSE, warning=FALSE }
datatable(readMappings,
  rownames = FALSE,
  options = list(
    pageLength = 10,
    lengthMenu = c(10, 20, 50)
  )
)
```
