---
title: "02 Options for CodelistGenerator(with full vocabulary)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{02b_Options_for_CodelistGenerator}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Set up 

```{r,  message=FALSE, warning=FALSE,echo=FALSE}
library(here)
library(readr)
library(DBI)
library(RSQLite)
library(here)
library(dplyr)
library(dbplyr)
library(stringr)
library(DT)
library(kableExtra)
library(CodelistGenerator)
```


## Creating a codelist for osteoarthritis
For this example we are going to generate a candidate codelist for osteoarthritis, looking at the impact of alternative search strategies.

## Set up 
```{r, eval=FALSE}
library(DBI)
library(RPostgres)
```

```{r, eval=FALSE}
# postgres database connection details
serverDbi <- Sys.getenv("server")
user <- Sys.getenv("user")
password <- Sys.getenv("password")
port <- Sys.getenv("port")
host <- Sys.getenv("host")

db <- dbConnect(RPostgres::Postgres(),
  dbname = serverDbi,
  port = port,
  host = host,
  user = user,
  password = password
)

# name of vocabulary schema
vocabularyDatabaseSchema <- "vocabulary"
```


## Search strategies
### Condition domain, without searching synonyms, without fuzzy match, with exclusions, without including descendants or ancestor
```{r,  message=FALSE, warning=FALSE,echo=FALSE}
oaCodes1 <- readRDS(here("vignettes", "optionsData01.RData"))
```

To start we will search for "osteoarthritis", while excluding "post-infection" and "post-traumatic", but without searching synonyms, without searching via source codes, without fuzzy matching, and without including descendants or the direct ancestor of the included concepts.

```{r, eval=FALSE}
oaCodes1 <- get_candidate_codes(
  keywords = "osteoarthritis",
  domains = "Condition",
  searchSynonyms = FALSE,
  searchSource = FALSE,
  fuzzyMatch = FALSE,
  exclude = c(
    "post-infection",
    "post-traumatic"
  ),
  includeDescendants = FALSE,
  includeAncestor = FALSE,
  db = db,
  vocabularyDatabaseSchema = vocabularyDatabaseSchema
)
```

What is the candidate codelist?
```{r,  message=FALSE, warning=FALSE }  
datatable(oaCodes1,
  options = list(
    pageLength = 10,
    lengthMenu = c(10, 20, 250)
  )
)
```


### Including descendants
```{r,  message=FALSE, warning=FALSE,echo=FALSE}
oaCodes2 <- readRDS(here("vignettes", "optionsData02.RData"))
```

Now we will also include the descendants of included concepts.
```{r, eval=FALSE}
oaCodes2 <- get_candidate_codes(
  keywords = "osteoarthritis",
  domains = "Condition",
  searchSynonyms = FALSE,
  searchSource = FALSE,
  fuzzyMatch = FALSE,
  exclude = c(
    "post-infection",
    "post-traumatic"
  ),
  includeDescendants = TRUE,
  includeAncestor = FALSE,
  db = db,
  vocabularyDatabaseSchema = vocabularyDatabaseSchema
)
```

What new codes do we pick up?
```{r,  message=FALSE, warning=FALSE }  
newCodes1To2 <- compareCodelists(oaCodes1, oaCodes2) %>%
  filter(codelist == "Only codelist 2") %>%
  select(-"codelist")

datatable(newCodes1To2,
  rownames = FALSE,
  options = list(
    pageLength = 10,
    lengthMenu = c(10, 20, 50)
  )
)
```

### Including observation domain
```{r,  message=FALSE, warning=FALSE,echo=FALSE}
oaCodes3 <- readRDS(here("vignettes", "optionsData03.RData"))
```

Now we will search the observation domain as well as the condition domain.
```{r, eval=FALSE}
oaCodes3 <- get_candidate_codes(
  keywords = "osteoarthritis",
  domains = c("Condition", "Observation"),
  searchSynonyms = FALSE,
  searchSource = FALSE,
  fuzzyMatch = FALSE,
  exclude = c(
    "post-infection",
    "post-traumatic"
  ),
  includeDescendants = FALSE,
  includeAncestor = FALSE,
  db = db,
  vocabularyDatabaseSchema = vocabularyDatabaseSchema
)
```

What new codes do we pick up?
```{r,  message=FALSE, warning=FALSE }  
newCodes1To3 <- compareCodelists(oaCodes1, oaCodes3) %>%
  filter(codelist == "Only codelist 2") %>%
  select(-"codelist")

datatable(newCodes1To3,
  rownames = FALSE,
  options = list(
    pageLength = 10,
    lengthMenu = c(10, 20, 50)
  )
)
```

### Search synonyms
```{r,  message=FALSE, warning=FALSE,echo=FALSE}
oaCodes4 <- readRDS(here("vignettes", "optionsData04.RData"))
```

Now we will search the concept synonym table to identify concepts to include.
```{r, eval=FALSE}
oaCodes4 <- get_candidate_codes(
  keywords = "osteoarthritis",
  domains = "Condition",
  searchSynonyms = TRUE,
  searchSource = FALSE,
  fuzzyMatch = FALSE,
  exclude = c(
    "post-infection",
    "post-traumatic"
  ),
  includeDescendants = FALSE,
  includeAncestor = FALSE,
  db = db,
  vocabularyDatabaseSchema = vocabularyDatabaseSchema
)
```

What new codes do we pick up?
```{r,  message=FALSE, warning=FALSE }  
newCodes1To4 <- compareCodelists(oaCodes1, oaCodes4) %>%
  filter(codelist == "Only codelist 2") %>%
  select(-"codelist")

datatable(newCodes1To4,
  rownames = FALSE,
  options = list(
    pageLength = 10,
    lengthMenu = c(10, 20, 50)
  )
)
```

### Search via source
```{r,  message=FALSE, warning=FALSE,echo=FALSE}
oaCodes5 <- readRDS(here("vignettes", "optionsData04.RData"))
```

Now we will search the concept synonym table to identify concepts to include.
```{r, eval=FALSE}
oaCodes5 <- get_candidate_codes(
  keywords = "osteoarthritis",
  domains = "Condition",
  searchSynonyms = FALSE,
  searchSource = TRUE,
  fuzzyMatch = FALSE,
  exclude = c(
    "post-infection",
    "post-traumatic"
  ),
  includeDescendants = FALSE,
  includeAncestor = FALSE,
  db = db,
  vocabularyDatabaseSchema = vocabularyDatabaseSchema
)
```

What new codes do we pick up?
```{r,  message=FALSE, warning=FALSE }  
newCodes1To5 <- compareCodelists(oaCodes1, oaCodes5) %>%
  filter(codelist == "Only codelist 2") %>%
  select(-"codelist")

datatable(newCodes1To5,
  rownames = FALSE,
  options = list(
    pageLength = 10,
    lengthMenu = c(10, 20, 50)
  )
)
```

### Using a fuzzy search
```{r,  message=FALSE, warning=FALSE,echo=FALSE}
oaCodes6 <- readRDS(here("vignettes", "optionsData05.RData"))
```

Now we will use an approximate search. See https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/agrep for further details (as this is what is running under the hood). Note, using the defaults of 0.1 for for substitutions, deletions, and insertions.
```{r, eval=FALSE}
oaCodes6 <- get_candidate_codes(
  keywords = "osteoarthritis",
  domains = "Condition",
  searchSynonyms = FALSE,
  searchSource = FALSE,
  fuzzyMatch = TRUE,
  maxDistanceCost = 0.1,
  exclude = c(
    "post-infection",
    "post-traumatic"
  ),
  includeDescendants = FALSE,
  includeAncestor = FALSE,
  db = db,
  vocabularyDatabaseSchema = vocabularyDatabaseSchema
)
```

What new codes do we pick up?
```{r,  message=FALSE, warning=FALSE }  
newCodes1To6 <- compareCodelists(oaCodes1, oaCodes6) %>%
  filter(codelist == "Only codelist 2") %>%
  select(-"codelist")

datatable(newCodes1To6,
  rownames = FALSE,
  options = list(
    pageLength = 10,
    lengthMenu = c(10, 20, 50)
  )
)
```

### Using a more generous fuzzy search
```{r,  message=FALSE, warning=FALSE,echo=FALSE}
oaCodes7 <- readRDS(here("vignettes", "optionsData06.RData"))
```

Here we widen the approximate searching, increasing the maximum distance allowed for a match to 0.2 for substitutions, deletions, and insertions.
```{r, eval=FALSE}
oaCodes7 <- get_candidate_codes(
  keywords = "osteoarthritis",
  domains = "Condition",
  searchSynonyms = FALSE,
  searchSource = FALSE,
  fuzzyMatch = TRUE,
  maxDistanceCost = 0.2,
  exclude = c(
    "post-infection",
    "post-traumatic"
  ),
  includeDescendants = FALSE,
  includeAncestor = FALSE,
  db = db,
  vocabularyDatabaseSchema = vocabularyDatabaseSchema
)
```

What new codes do we pick up?
```{r,  message=FALSE, warning=FALSE }  
newCodes1To7 <- compareCodelists(oaCodes1, oaCodes7) %>%
  filter(codelist == "Only codelist 2") %>%
  select(-"codelist")

datatable(newCodes1To7,
  rownames = FALSE,
  options = list(
    pageLength = 10,
    lengthMenu = c(10, 20, 50)
  )
)
```

### Include ancestor
```{r,  message=FALSE, warning=FALSE,echo=FALSE}
oaCodes8 <- readRDS(here("vignettes", "optionsData07.RData"))
```

Now we include the direct ancestor of included terms.
```{r, eval=FALSE}
oaCodes8 <- get_candidate_codes(
  keywords = "osteoarthritis",
  domains = "Condition",
  searchSynonyms = FALSE,
  searchSource = FALSE,
  fuzzyMatch = FALSE,
  maxDistanceCost = 0.1,
  exclude = c(
    "post-infection",
    "post-traumatic"
  ),
  includeDescendants = FALSE,
  includeAncestor = TRUE,
  db = db,
  vocabularyDatabaseSchema = vocabularyDatabaseSchema
)
```

What new codes do we pick up?
```{r,  message=FALSE, warning=FALSE }  
newCodes1To8 <- compareCodelists(oaCodes1, oaCodes8) %>%
  filter(codelist == "Only codelist 2") %>%
  select(-"codelist")

datatable(newCodes1To8,
  rownames = FALSE,
  options = list(
    pageLength = 10,
    lengthMenu = c(10, 20, 50)
  )
)
```
