---
title: "04 Codelists for medications"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{a04_codelists_for_medications}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r,  message=FALSE, warning=FALSE,echo=FALSE}
library(here)
library(readr)
library(DBI)
library(RSQLite)
library(here)
library(dplyr)
library(stringr)
library(DT)
library(kableExtra)
library(CodelistGenerator)
```

## Creating a codelists for colonoscopy
For this vignette we are going to create codelists for identifying a colonoscopy. For this example we are going to generate a candidate codelists looking for codes in either the procedure domain.

## Connecting to a database with the OMOP CDM vocabularies
As described previously there are alternative ways to link to the vocabularies. A connection to an existing database can be made like so:

```{r, eval=FALSE}
library(DBI)
library(RPostgres)
```

```{r, eval=FALSE}
# postgres database connection details
server_dbi<-Sys.getenv("server")
user<-Sys.getenv("user")
password<- Sys.getenv("password")
port<-Sys.getenv("port")
host<-Sys.getenv("host")

db <- dbConnect(RPostgres::Postgres(),
                dbname = server_dbi,
                port = port,
                host = host,
                user = user,
                password = password)

# name of vocabulary schema
vocabulary_database_schema<-"vocabulary"
```

## Codelist from "colonoscopy" and its descendants
First we can follow the approach of identifying high-level codes and include all their descendants.

```{r, eval=FALSE}
library(dplyr)
library(stringr)
library(DT)
library(kableExtra)
```

```{r, eval=FALSE}
codes_from_descendants<-tbl(db,
  sql(paste0("SELECT * FROM ",
     vocabulary_database_schema,
     ".concept_ancestor"))) %>%
  filter(ancestor_concept_id %in% c("1503297")) %>%
  select("descendant_concept_id") %>%
  rename("concept_id"="descendant_concept_id") %>%
  inner_join(tbl(db, sql(paste0("SELECT * FROM ",
     vocabulary_database_schema,
     ".concept"))))%>%
  select("concept_id", "concept_name",
         "domain_id", "vocabulary_id") %>%
  collect()

```

```{r,  message=FALSE, warning=FALSE,echo=FALSE}
codes_from_descendants<-readRDS(here("vignettes","med_data_01.RData"))
```

```{r,  message=FALSE, warning=FALSE }  
datatable(codes_from_descendants,
          rownames=FALSE,
          options = list(pageLength = 10,
                         lengthMenu = c(10, 20,50)))

```

## Generating a candidate codelist using Codelist Generator
But we can also use CodelistGenerator to generate a candidate codelist for colonoscopy.

```{r, eval=FALSE }
library(CodelistGenerator)
```

Let's do a search for a single keyword of "colonoscopy" without any exclusions and not including descendants
```{r, eval=FALSE }
metformin_codes2<-get_candidate_codes(keywords="metformin",
                    domains=c("Drug"),
                    search_synonyms = FALSE,
                    fuzzy_match = FALSE,
                    exclude = NULL,
                    include_descendants = TRUE,
                    include_ancestor = FALSE,
                    verbose = TRUE ,
                    db=db,
                    vocabulary_database_schema =  vocabulary_database_schema)

```

```{r,  message=FALSE, warning=FALSE,echo=FALSE}
metformin_codes2<-readRDS(here("vignettes","metformin_codes2.RData"))
```

```{r,  message=FALSE, warning=FALSE }
datatable(metformin_codes2,
          rownames=FALSE,
          options = list(
  pageLength = 10,
  lengthMenu = c(10, 20,50)
))
```

What is the difference between this codelist and the one based on descendants?
```{r,  eval=TRUE }
code_comparison<-full_join(codes_from_descendants  %>% mutate(type1="Specified codes and descendants"), 
                     metformin_codes2  %>% mutate(type2="CodelistGenerator")) %>% 
  mutate(type=ifelse(!is.na(type1) & !is.na(type2),
                     "Both",
              ifelse(!is.na(type1) & is.na(type2),
                    "Only specified codes and descendants",
              ifelse(is.na(type1) & !is.na(type2),
                     "Only CodelistGenerator",
                     NA)))) %>% 
  select(-c("type1","type2"))

```


```{r,  message=FALSE, warning=FALSE } 
kable(code_comparison %>% 
  group_by(type) %>% 
  tally())
```

What are these extra codes picked up by CodelistGenerator?
```{r,  message=FALSE, warning=FALSE }
datatable(code_comparison %>% 
            filter(type=="Only CodelistGenerator"),
          rownames=FALSE,
          options = list(
  pageLength = 10,
  lengthMenu = c(10, 20,50)
))
```
