---
title: "04 Codelists for medications"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{a04_codelists_for_medications}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r,  message=FALSE, warning=FALSE,echo=FALSE}
library(here)
library(readr)
library(DBI)
library(RSQLite)
library(here)
library(dplyr)
library(dbplyr)
library(stringr)
library(DT)
library(kableExtra)
library(CodelistGenerator)
```

## Creating a codelists for metformin
For this vignette we are going to create codelists for identifying a metformin, looking in the drug domain.

## Connecting to a database with the OMOP CDM vocabularies
```{r, eval=FALSE}
library(DBI)
library(RPostgres)
```

```{r, eval=FALSE}
# postgres database connection details
serverDbi <- Sys.getenv("server")
user <- Sys.getenv("user")
password <- Sys.getenv("password")
port <- Sys.getenv("port")
host <- Sys.getenv("host")

db <- dbConnect(RPostgres::Postgres(),
  dbname = serverDbi,
  port = port,
  host = host,
  user = user,
  password = password
)

# name of vocabulary schema
vocabularyDatabaseSchema <- "vocabulary"
```

## Codelist from "metformin" and its descendants
First we can follow the approach of identifying high-level codes and include all their descendants.

```{r, eval=FALSE}
library(dplyr)
library(dbplyr)
library(stringr)
library(DT)
library(kableExtra)
```

```{r, eval=FALSE}
codesFromDescendants <- tbl(
  db,
  sql(paste0(
    "SELECT * FROM ",
    vocabularyDatabaseSchema,
    ".concept_ancestor"
  ))
) %>%
  filter(ancestor_concept_id %in% c("1503297")) %>%
  select("descendant_concept_id") %>%
  rename("concept_id" = "descendant_concept_id") %>%
  inner_join(tbl(db, sql(paste0(
    "SELECT * FROM ",
    vocabularyDatabaseSchema,
    ".concept"
  )))) %>%
  select(
    "concept_id", "concept_name",
    "domain_id", "vocabulary_id"
  ) %>%
  collect()
```

```{r,  message=FALSE, warning=FALSE,echo=FALSE}
codesFromDescendants <- readRDS(here("vignettes", "medData01.RData"))
```

```{r,  message=FALSE, warning=FALSE }  
datatable(codesFromDescendants,
  rownames = FALSE,
  options = list(
    pageLength = 10,
    lengthMenu = c(10, 20, 50)
  )
)
```

## Generating a candidate codelist using Codelist Generator
But we can also use CodelistGenerator to generate a candidate codelist for metformin,

```{r, eval=FALSE }
library(CodelistGenerator)
```

Let's do a search for a single keyword of "metformin", including descendants.
```{r, eval=FALSE }
metforminCodes2 <- getCandidateCodes(
  keywords = "metformin",
  domains = c("Drug"),
  includeDescendants = TRUE,
  db = db,
  vocabularyDatabaseSchema = vocabularyDatabaseSchema
)
```

```{r,  message=FALSE, warning=FALSE,echo=FALSE}
metforminCodes2 <- readRDS(here("vignettes", "metforminCodes2.RData"))
```

```{r,  message=FALSE, warning=FALSE }
datatable(metforminCodes2,
  rownames = FALSE,
  options = list(
    pageLength = 10,
    lengthMenu = c(10, 20, 50)
  )
)
```

What is the difference between this codelist and the one based on descendants?
```{r,  eval=TRUE }
codeComparison <- compareCodelists(codesFromDescendants, metforminCodes2)
```


```{r,  message=FALSE, warning=FALSE } 
kable(codeComparison %>%
  group_by(codelist) %>%
  tally())
```

What are these extra codes picked up by CodelistGenerator?
```{r,  message=FALSE, warning=FALSE }
datatable(codeComparison %>%
  filter(codelist == "Only codelist 2"),
rownames = FALSE,
options = list(
  pageLength = 10,
  lengthMenu = c(10, 20, 50)
)
)
```
