---
title: "Code lists for medications"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{a04_codelists_for_medications}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r,  message=FALSE, warning=FALSE,echo=FALSE}
library(here)
library(readr)
library(DBI)
library(here)
library(dplyr)
library(dbplyr)
library(stringr)
library(DT)
library(kableExtra)
library(CodelistGenerator)
```

## Creating a codelists for acetaminophen
For this vignette we are going to create codelists for identifying a acetaminophen, looking in the drug domain.

## Connecting to a database with the OMOP CDM vocabularies
```{r, eval=FALSE}
# postgres database connection details
serverDbi <- Sys.getenv("server")
user <- Sys.getenv("user")
password <- Sys.getenv("password")
port <- Sys.getenv("port")
host <- Sys.getenv("host")

db <- DBI::dbConnect(RPostgres::Postgres(),
  dbname = serverDbi,
  port = port,
  host = host,
  user = user,
  password = password
)

# name of vocabulary schema
vocabularyDatabaseSchema <- "vocabulary"
```

## Codelist from "acetaminophen" and its descendants
First we can follow the approach of identifying high-level codes and include all their descendants.

```{r, eval=FALSE}
library(dplyr)
library(CodelistGenerator)
library(stringr)
library(DT)
library(kableExtra)
```

```{r, eval=FALSE}
acetaminophen1 <- getCandidateCodes(
  cdm = cdm,
  keywords = "acetaminophen",
  domains = "drug",
  standardConcept = "standard",
  includeDescendants = TRUE
)
```

```{r,  message=FALSE, warning=FALSE, echo=FALSE}
acetaminophen1 <- readRDS(here("vignettes", "medData01.RData"))
```

```{r,  message=FALSE, warning=FALSE }  
acetaminophen1 %>% dplyr::glimpse()
```

## Restrict on dose form
We can also restrict on dose form, in this case requiring dose form to include either "injection" or "intravenous" in it's name
```{r, eval=FALSE}
acetaminophen2 <- getCandidateCodes(
  cdm = cdm_arrow,
  keywords = "acetaminophen",
  domains = "drug",
  doseForm = c("injection", "intravenous"),
  standardConcept = "standard",
  includeDescendants = TRUE,
  verbose = TRUE
)
```

```{r,  message=FALSE, warning=FALSE, echo=FALSE}
acetaminophen2 <- readRDS(here("vignettes", "medData02.RData"))
```

```{r,  message=FALSE, warning=FALSE }  
acetaminophen2 %>% dplyr::glimpse()
```

## Restrict on concept class
We can also restrict on concept class, in this case only identifying codes with a concept class of "Quant Clinical Drug"

```{r, eval=FALSE}
acetaminophen3 <- getCandidateCodes(
  cdm = cdm_arrow,
  keywords = "acetaminophen",
  domains = "drug",
  conceptClassId = c("Quant Clinical Drug"),
  doseForm = c("injection", "intravenous"),
  standardConcept = "standard",
  includeDescendants = TRUE,
  verbose = TRUE
)
```

```{r,  message=FALSE, warning=FALSE, echo=FALSE}
acetaminophen3 <- readRDS(here("vignettes", "medData03.RData"))
```

```{r,  message=FALSE, warning=FALSE }  
acetaminophen3 %>% dplyr::glimpse()
```
