---
title: "Code lists for medications"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{a04_codelists_for_medications}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r,  message=FALSE, warning=FALSE,echo=FALSE}
library(here)
library(readr)
library(DBI)
library(here)
library(dplyr)
library(dbplyr)
library(stringr)
library(DT)
library(kableExtra)
library(CodelistGenerator)
```

## Creating a codelists for acetaminophen
For this vignette we are going to create codelists for identifying a acetaminophen, looking in the drug domain.

## Connecting to a database with the OMOP CDM vocabularies
```{r, eval=FALSE}
# postgres database connection details
serverDbi <- Sys.getenv("server")
user <- Sys.getenv("user")
password <- Sys.getenv("password")
port <- Sys.getenv("port")
host <- Sys.getenv("host")

db <- DBI::dbConnect(RPostgres::Postgres(),
  dbname = serverDbi,
  port = port,
  host = host,
  user = user,
  password = password
)

# name of vocabulary schema
vocabularyDatabaseSchema <- "vocabulary"
```

## Codelist from "acetaminophen" and its descendants
First we can follow the approach of identifying high-level codes and include all their descendants.

```{r, eval=FALSE}
library(dplyr)
library(CodelistGenerator)
library(stringr)
library(DT)
library(kableExtra)
```

```{r, eval=FALSE}
acetaminophen1 <- getCandidateCodes(
  cdm = cdm,
  keywords = "acetaminophen",
  domains = "drug",
  standardConcept = "standard",
  includeDescendants = TRUE
)
```

```{r,  message=FALSE, warning=FALSE, echo=FALSE}
acetaminophen1 <- readRDS(here("vignettes", "medData01.RData"))
```

```{r,  message=FALSE, warning=FALSE }  
acetaminophen1 %>% dplyr::glimpse()
```

## Restrict on dose form
We can also restrict on dose form. In this case we can specify "injection" or "intravenous" either in the keywords or by using the dose form option.

```{r, eval=FALSE}
acetaminophen2a <- getCandidateCodes(
  cdm = cdm,
  keywords = c(
    "acetaminophen injection",
    "acetaminophen intravenous"
  ),
  domains = "drug",
  standardConcept = "standard",
  includeDescendants = TRUE,
  verbose = TRUE
)

acetaminophen2b <- getCandidateCodes(
  cdm = cdm,
  keywords = "acetaminophen",
  domains = "drug",
  doseForm = c("injection", "intravenous"),
  standardConcept = "standard",
  includeDescendants = TRUE,
  verbose = TRUE
)
```

```{r,  message=FALSE, warning=FALSE, echo=FALSE}
acetaminophen2a <- readRDS(here("vignettes", "medData02a.RData"))
acetaminophen2b <- readRDS(here("vignettes", "medData02b.RData"))
```

```{r,  message=FALSE, warning=FALSE }  
acetaminophen2a %>% dplyr::glimpse()
acetaminophen2b %>% dplyr::glimpse()
```

In this case the two searches return the same result.
```{r,  message=FALSE, warning=FALSE}
compareCodelists(acetaminophen2a, acetaminophen2b) %>% 
  group_by(codelist) %>% 
  tally()
```

We can see all the dose forms that could be specified like so
```{r, eval=FALSE}
getDoseForm(cdm = cdm)
```

```{r,  message=FALSE, warning=FALSE, echo=FALSE}
readRDS(here("vignettes", "medDataDoseForms.RData"))
```

## Restrict on concept class
We can also restrict on concept class, in this case only identifying codes with a concept class of "Quant Clinical Drug"

```{r, eval=FALSE}
acetaminophen3 <- getCandidateCodes(
  cdm = cdm,
  keywords = "acetaminophen",
  domains = "drug",
  conceptClassId = c("Quant Clinical Drug"),
  doseForm = c("injection", "intravenous"),
  standardConcept = "standard",
  includeDescendants = TRUE,
  verbose = TRUE
)
```

```{r,  message=FALSE, warning=FALSE, echo=FALSE}
acetaminophen3 <- readRDS(here("vignettes", "medData03.RData"))
```

```{r,  message=FALSE, warning=FALSE }  
acetaminophen3 %>% dplyr::glimpse()
```

We can see all the concept classes for drugs like so
```{r, eval=FALSE}
getConceptClassId(cdm = cdm, domain = "drug")
```

```{r,  message=FALSE, warning=FALSE, echo=FALSE}
readRDS(here("vignettes", "medDataConceptClass.RData"))
```
