---
title: "Introduction to CodelistGenerator"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction_to_CodelistGenerator}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r,  message=FALSE, warning=FALSE,echo=FALSE}
library(here)
library(readr)
library(DBI)
library(RSQLite)
library(here)
library(dplyr)
library(stringr)
library(DT)
library(kableExtra)
library(CodelistGenerator)
```


## Creating a codelist for dementia
For this example we are going to generate a candidate codelist for dementia, only looking for codes in the condition domain.

## Getting the OMOP CDM vocabularies
If you do not have ready access to a database with data in the OMOP CDM format (which will necessarily contain the vocabulary tables), you will first need to obtain the OMOP CDM vocabularies from https://athena.ohdsi.org. Once these are downloaded, we can make a vocabulary only database like so:

```{r, eval=FALSE}
library(readr)
library(DBI)
library(RSQLite)
library(here)
```

```{r, eval=FALSE}
vocab.folder<-Sys.getenv("omop_cdm_vocab_path") # path to directory of unzipped files
concept<-read_delim(paste0(vocab.folder,"/CONCEPT.csv"),
     "\t", escape_double = FALSE, trim_ws = TRUE)
concept_relationship<-read_delim(paste0(vocab.folder,"/CONCEPT_RELATIONSHIP.csv"),
     "\t", escape_double = FALSE, trim_ws = TRUE) 
concept_ancestor<-read_delim(paste0(vocab.folder,"/CONCEPT_ANCESTOR.csv"),
     "\t", escape_double = FALSE, trim_ws = TRUE)
concept_synonym<-read_delim(paste0(vocab.folder,"/CONCEPT_SYNONYM.csv"),
     "\t", escape_double = FALSE, trim_ws = TRUE)
vocabulary<-read_delim(paste0(vocab.folder,"/VOCABULARY.csv"),
     "\t", escape_double = FALSE, trim_ws = TRUE)

db <- dbConnect(RSQLite::SQLite(), ":memory:")
dbWriteTable(db, "concept", concept, overwrite=TRUE)
dbWriteTable(db, "concept_relationship", concept_relationship, overwrite=TRUE)
dbWriteTable(db, "concept_ancestor", concept_ancestor, overwrite=TRUE)
dbWriteTable(db, "concept_synonym", concept_synonym, overwrite=TRUE)
dbWriteTable(db, "vocabulary", vocabulary)
rm(concept,concept_relationship, concept_ancestor, concept_synonym, vocabulary)
vocabulary_database_schema<-"main"
```

The structure of each of these tables is described in detail at: https://ohdsi.github.io/CommonDataModel/cdm53.html#Vocabulary_Tables 

## Check version of the vocabularies
It is important to note that the results from CodelistGenerator will be specific to a particular version of the OMOP CDM vocabularies. We can see the version of the vocabulary being used like so
```{r, eval=FALSE}
get_vocab_version(db=db,
                  vocabulary_database_schema = "main")
```

```{r,  message=FALSE, warning=FALSE,echo=FALSE}
vocab_version<-readRDS(here("vignettes","intro_vocab.RData"))
vocab_version
```


## Codelist from "Dementia" (4182210) and its descendants
The simplest approach to identifying potential codes is to take a high-level code and include all its descendants. 
```{r, eval=FALSE}
library(dplyr)
library(stringr)
library(DT)
library(kableExtra)
```

```{r, eval=FALSE}
codes_from_descendants<-tbl(db, 
  sql(paste0("SELECT * FROM ",
     vocabulary_database_schema,
     ".concept_ancestor"))) %>% 
  filter(ancestor_concept_id=="4182210") %>% 
  select("descendant_concept_id") %>% 
  rename("concept_id"="descendant_concept_id") %>% 
  left_join(tbl(db, sql(paste0("SELECT * FROM ",
     vocabulary_database_schema,
     ".concept"))))%>%
  select("concept_id", "concept_name", 
         "domain_id", "vocabulary_id") %>% 
  collect()
```

```{r,  message=FALSE, warning=FALSE,echo=FALSE}
codes_from_descendants<-readRDS(here("vignettes","intro_data_01.RData"))
```

```{r,  message=FALSE, warning=FALSE }  
datatable(codes_from_descendants,
          rownames=FALSE,
          options = list(pageLength = 10,
                         lengthMenu = c(10, 20,50)))

```

This looks to pick up most relevant codes. But, this approach misses codes that are not a descendant of 4182210. For example, codes such as "Wandering due to dementia" (37312577; https://athena.ohdsi.org/search-terms/terms/37312577) and "Anxiety due to dementia" (37312031; https://athena.ohdsi.org/search-terms/terms/37312031) are not picked up.

## Generating a candidate codelist using Codelist Generator
To try and include all such terms that could be included we can use CodelistGenerator.

```{r, eval=FALSE }
library(CodelistGenerator)
```

First, let's do a simple search for a single keyword of "dementia" without any exclusions, without searching synonyms, and including descendants but not ancestors.
```{r, eval=FALSE }
dementia_codes1<-get_candidate_codes(keywords="dementia",
                    domains="Condition",
                    search_synonyms = FALSE,
                    fuzzy_match = FALSE,
                    exclude = NULL,
                    include_descendants = TRUE,
                    include_ancestor = FALSE,
                    db=db,
                    vocabulary_database_schema =  vocabulary_database_schema)
```

```{r,  message=FALSE, warning=FALSE,echo=FALSE}
dementia_codes1<-readRDS(here("vignettes","intro_data_02.RData"))
```

```{r,  message=FALSE, warning=FALSE }
datatable(dementia_codes1,
          rownames=FALSE,
          options = list(
  pageLength = 10,
  lengthMenu = c(10, 20,50)
))
```

What is the difference between this codelist and the one from 4182210 and its descendants?
```{r,  eval=FALSE }
code_comparison<-full_join(codes_from_descendants  %>% mutate(type1="4182210 and descendants"), 
                     dementia_codes1  %>% mutate(type2="CodelistGenerator")) %>% 
  mutate(type=ifelse(!is.na(type1) & !is.na(type2),
                     "Both",
              ifelse(!is.na(type1) & is.na(type2),
                     "Only 4182210 and descendants",
              ifelse(is.na(type1) & !is.na(type2),
                     "Only CodelistGenerator",
                     NA)))) %>% 
  select(-c("type1","type2"))
```

```{r,  message=FALSE, warning=FALSE,echo=FALSE}
code_comparison<-readRDS(here("vignettes","intro_data_03.RData"))
```

```{r,  message=FALSE, warning=FALSE } 
kable(code_comparison %>% 
  group_by(type) %>% 
  tally())
```

What are these extra codes picked up by CodelistGenerator?
```{r,  message=FALSE, warning=FALSE }
datatable(code_comparison %>% 
            filter(type=="Only CodelistGenerator"),
          rownames=FALSE,
          options = list(
  pageLength = 10,
  lengthMenu = c(10, 20,50)
))
```

## Review mappings from source vocabularies
Perhaps we want to see what ICD10CM codes map to our candidate codelist. We can get these by running

```{r,  message=FALSE, warning=FALSE,echo=FALSE}
icd_mappings<-readRDS(here("vignettes","intro_data_04.RData"))
```

```{r,  eval=FALSE }
icd_mappings<-show_mappings(candidate_codelist=dementia_codes4,
                    source_vocabularies="ICD10CM",
                    db=db,
                    vocabulary_database_schema =  vocabulary_database_schema)
```

```{r,  message=FALSE, warning=FALSE }
datatable(icd_mappings,
          rownames=FALSE,
          options = list(
  pageLength = 10,
  lengthMenu = c(10, 20,50)
))
```


```{r,  message=FALSE, warning=FALSE,echo=FALSE}
read_mappings<-readRDS(here("vignettes","intro_data_05.RData"))
```

```{r,  eval=FALSE }
read_mappings<-show_mappings(candidate_codelist=dementia_codes4,
                    source_vocabularies="Read",
                    db=db,
                    vocabulary_database_schema =  vocabulary_database_schema)
```

```{r,  message=FALSE, warning=FALSE }
datatable(read_mappings,
          rownames=FALSE,
          options = list(
  pageLength = 10,
  lengthMenu = c(10, 20,50)
))
```

