---
title: "Options for CodelistGenerator"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Options_for_CodelistGenerator}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r,  message=FALSE, warning=FALSE,echo=FALSE}
library(here)
library(readr)
library(DBI)
library(RSQLite)
library(here)
library(dplyr)
library(stringr)
library(DT)
library(kableExtra)
library(CodelistGenerator)
```


## Creating a codelist for osteoarthritis
For this example we are going to generate a candidate codelist for osteoarthritis. 

## Set up 
```{r, eval=FALSE}
library(readr)
library(DBI)
library(RSQLite)
library(here)
```

```{r, eval=FALSE}
# postgres database connection details
server_dbi<-Sys.getenv("server")
user<-Sys.getenv("user")
password<- Sys.getenv("password")
port<-Sys.getenv("port")
host<-Sys.getenv("host")

db <- dbConnect(RPostgres::Postgres(),
                dbname = server_dbi,
                port = port,
                host = host,
                user = user,
                password = password)

# name of vocabulary schema
vocabulary_database_schema<-"vocabulary"
```


## Search strategies
### Condition domain, without searching synonyms, without fuzzy match, with exclusions, without including descendants or ancestor
```{r,  message=FALSE, warning=FALSE,echo=FALSE}
oa_codes1<-readRDS(here("vignettes","options_data_01.RData"))
```

To start we will search for "osteoarthritis", while excluding "post-infection" and "post-traumatic", but without searching synonyms, without fuzzy match, and without including descendants or the direct ancestor of the included concepts.

```{r, eval=FALSE}
oa_codes1<-get_candidate_codes(keywords="osteoarthritis",
                    domains="Condition",
                    search_synonyms = FALSE,
                    fuzzy_match = FALSE,
                    exclude = c("post-infection",
                                "post-traumatic"),
                    include_descendants = FALSE,
                    include_ancestor = FALSE,
                    db=db,
                    vocabulary_database_schema =  vocabulary_database_schema)
```

What is the candidate codelist?
```{r,  message=FALSE, warning=FALSE }  
datatable(oa_codes1,
          options = list(pageLength = 10,
                         lengthMenu = c(10, 20,250)))

```


### Including descendants
```{r,  message=FALSE, warning=FALSE,echo=FALSE}
oa_codes2<-readRDS(here("vignettes","options_data_02.RData"))
```

Now we will also include the descendants of included concepts.
```{r, eval=FALSE}
oa_codes2<-get_candidate_codes(keywords="osteoarthritis",
                    domains="Condition",
                    search_synonyms = FALSE,
                    fuzzy_match = FALSE,
                    exclude = c("post-infection",
                                "post-traumatic"),
                    include_descendants = TRUE,
                    include_ancestor = FALSE,
                    db=db,
                    vocabulary_database_schema =  vocabulary_database_schema)
```

What new codes do we pick up?
```{r,  message=FALSE, warning=FALSE }  
new_codes_1_2<-bind_rows(oa_codes1, oa_codes2) %>% 
  group_by(concept_id) %>% 
  mutate(seq=length(concept_name)) %>%
  filter(seq==1)
datatable(new_codes_1_2,
          rownames=FALSE,
          options = list(
  pageLength = 10,
  lengthMenu = c(10, 20,50)
))

```

### Including observation domain
```{r,  message=FALSE, warning=FALSE,echo=FALSE}
oa_codes3<-readRDS(here("vignettes","options_data_03.RData"))
```

Now we will search the observation domain as well as the condition domain.
```{r, eval=FALSE}
oa_codes3<-get_candidate_codes(keywords="osteoarthritis",
                    domains=c("Condition","Observation"),
                    search_synonyms = FALSE,
                    fuzzy_match = FALSE,
                    max_distance_substitutions = 0.1,
                    max_distance_deletions = 0.1,
                    max_distance_insertions = 0.1,
                    exclude = c("post-infection",
                                "post-traumatic"),
                    include_descendants = FALSE,
                    include_ancestor = FALSE,
                    db=db,
                    vocabulary_database_schema =  vocabulary_database_schema)
```

What new codes do we pick up?
```{r,  message=FALSE, warning=FALSE }  
new_codes_1_3<-bind_rows(oa_codes1, oa_codes3) %>% 
  group_by(concept_id) %>% 
  mutate(seq=length(concept_name)) %>%
  filter(seq==1)
datatable(new_codes_1_3,
          rownames=FALSE,
          options = list(
  pageLength = 10,
  lengthMenu = c(10, 20,50)
))
```

### Search synonyms
```{r,  message=FALSE, warning=FALSE,echo=FALSE}
oa_codes4<-readRDS(here("vignettes","options_data_04.RData"))
```

Now we will search the concept synonym table to identify concepts to include.
```{r, eval=FALSE}
oa_codes4<-get_candidate_codes(keywords="osteoarthritis",
                    domains="Condition",
                    search_synonyms = TRUE,
                    fuzzy_match = FALSE,
                    max_distance_substitutions = 0.1,
                    max_distance_deletions = 0.1,
                    max_distance_insertions = 0.1,
                    exclude = c("post-infection",
                                "post-traumatic"),
                    include_descendants = FALSE,
                    include_ancestor = FALSE,
                    db=db,
                    vocabulary_database_schema =  vocabulary_database_schema)
```

What new codes do we pick up?
```{r,  message=FALSE, warning=FALSE }  
new_codes_1_4<-bind_rows(oa_codes1, oa_codes4) %>% 
  group_by(concept_id) %>% 
  mutate(seq=length(concept_name)) %>%
  filter(seq==1)
datatable(new_codes_1_4,
          rownames=FALSE,
          options = list(
  pageLength = 10,
  lengthMenu = c(10, 20,50)
))
```

### Using a fuzzy search
```{r,  message=FALSE, warning=FALSE,echo=FALSE}
oa_codes5<-readRDS(here("vignettes","options_data_05.RData"))
```

Now we will use an approximate search. See https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/agrep for further details (as this is what is running under the hood). Note, we set the default max distance as 0.1 (equivalent to 10% of the pattern length).
```{r, eval=FALSE}
oa_codes5<-get_candidate_codes(keywords="osteoarthritis",
                    domains="Condition",
                    search_synonyms = FALSE,
                    fuzzy_match = TRUE,
                    max_distance_substitutions = 0.1,
                    max_distance_deletions = 0.1,
                    max_distance_insertions = 0.1,
                    exclude = c("post-infection",
                                "post-traumatic"),
                    include_descendants = FALSE,
                    include_ancestor = FALSE,
                    db=db,
                    vocabulary_database_schema =  vocabulary_database_schema)
```

What new codes do we pick up?
```{r,  message=FALSE, warning=FALSE }  
new_codes_1_5<-bind_rows(oa_codes1, oa_codes5) %>% 
  group_by(concept_id) %>% 
  mutate(seq=length(concept_name)) %>%
  filter(seq==1)
datatable(new_codes_1_5,
          rownames=FALSE,
          options = list(
  pageLength = 10,
  lengthMenu = c(10, 20,50)
))
```

### Using a more generous fuzzy search
```{r,  message=FALSE, warning=FALSE,echo=FALSE}
oa_codes6<-readRDS(here("vignettes","options_data_06.RData"))
```

Here we widen the approximate searching, increasing the maximum distance allowed for a match to 0.2 (equivalent to 20% of the pattern length).
```{r, eval=FALSE}
oa_codes6<-get_candidate_codes(keywords="osteoarthritis",
                    domains="Condition",
                    search_synonyms = FALSE,
                    fuzzy_match = TRUE,
                    max_distance_substitutions = 0.2,
                    max_distance_deletions = 0.2,
                    max_distance_insertions = 0.2,
                    exclude = c("post-infection",
                                "post-traumatic"),
                    include_descendants = FALSE,
                    include_ancestor = FALSE,
                    db=db,
                    vocabulary_database_schema =  vocabulary_database_schema)
```

What new codes do we pick up?
```{r,  message=FALSE, warning=FALSE }  
new_codes_1_6<-bind_rows(oa_codes1, oa_codes6) %>% 
  group_by(concept_id) %>% 
  mutate(seq=length(concept_name)) %>%
  filter(seq==1)
datatable(new_codes_1_6,
          rownames=FALSE,
          options = list(
  pageLength = 10,
  lengthMenu = c(10, 20,50)
))
```

### Include ancestor
```{r,  message=FALSE, warning=FALSE,echo=FALSE}
oa_codes7<-readRDS(here("vignettes","options_data_07.RData"))
```

Now we include the direct ancestor of included terms.
```{r, eval=FALSE}
oa_codes7<-get_candidate_codes(keywords="osteoarthritis",
                    domains="Condition",
                    search_synonyms = FALSE,
                    fuzzy_match = FALSE,
                    max_distance_substitutions = 0.1,
                    max_distance_deletions = 0.1,
                    max_distance_insertions = 0.1,
                    exclude = c("post-infection",
                                "post-traumatic"),
                    include_descendants = FALSE,
                    include_ancestor = TRUE,
                    db=db,
                    vocabulary_database_schema =  vocabulary_database_schema)
```

What new codes do we pick up?
```{r,  message=FALSE, warning=FALSE }  
new_codes_1_7<-bind_rows(oa_codes1, oa_codes7) %>% 
  group_by(concept_id) %>% 
  mutate(seq=length(concept_name)) %>%
  filter(seq==1)
datatable(new_codes_1_7,
          rownames=FALSE,
          options = list(
  pageLength = 10,
  lengthMenu = c(10, 20,50)
))
```

